# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data
Delete intermediate files
```{r}
url <- 'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'
zipfile <- 'activity.zip'
download.file(url, method="curl", destfile=zipfile)
unzip(zipfile)
activity <- read.csv('activity.csv')
file.remove('activity.csv')
str(activity)
```



## What is mean total number of steps taken per day
```{r}
hist(activity$steps)
summary(activity$steps)
``` 



## What is the average daily activity pattern?
```{r}
library(ggplot2)
ggplot(activity, aes(interval, steps)) + geom_line(stat='summary', fun.y='mean') + theme_minimal()

library(dplyr)

steps.mean <- activity %.% 
  select(interval, steps) %.%
  group_by(interval) %.%
  summarise(meansteps = mean(steps,na.rm=TRUE)) 

with(steps.mean, plot(interval, meansteps, type = 'l'))
```

Which interval
```{r}
with(steps.mean, interval[which(max(meansteps) == meansteps)])
```
which (pun intended) matches with the chart

## Imputing missing values
complete.cases counts number of rows without any NA
the number of rows - complete.cases gives us the number of incomplete rows

Use the mean step per date/
```{r}
nrow(activity) - sum(complete.cases(activity))
library(Hmisc)
```


Impute the activity dataset with the means steps per interval (allready calculated)
merge the activity dataframe with the means per interval
```{r}
temp <- merge(activity, steps.mean)
temp[is.na(temp$steps),'steps'] <- temp$meansteps[is.na(temp$steps)]
```

Create plots comparing original data with imputed data. Imputed data have as expected as higher
number of steps.
```{r}
par(mfrow = c(2,1), mar = c(4,4,1,1))
hist(activity$steps)
hist(temp$steps)
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(lubridate)
activity$date <- as.Date(activity$date)
activity$weekdayorweekend <- ifelse(wday(activity$date) %in% c(1,7), 'Weekend', 'Weekday')

activity.daytype <- 
  activity %.% group_by(interval, weekdayorweekend) %.% summarise(steps = mean(steps,na.rm=TRUE))


xyplot(steps ~ interval | weekdayorweekend,  activity.daytype, type="l")
ggplot(activity.daytype, aes(interval, steps)) + geom_line() + facet_wrap(~weekdayorweekend, ncol=1)

```


